---
title: "plastic-waste-step-2"
format: html
---

# Plastic Waste - Step 2

```{r}
library(tidyverse)
```

## Reading in the Data
```{r}
plastics <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2021/2021-01-26/plastics.csv')
```

## Proportion of Polyester Plastic Between 2019 & 2020 By Country
*New Column: proportion of Polyester Plastics for each year*

*Uses iteration, no custom function*
```{r}
pet_props <- plastics |> 
  group_by(year, country) |>
  summarize(
   prop = sum(pet, na.rm = TRUE)/sum(grand_total, na.rm = TRUE)
   )
pet_props
```

*Custom Function*

```{r}
#| echo: true
#| output: false
Yearly_PET_Prop <- function(input_country){
  pet_props |>
    filter(country == input_country) 
}

countries <- plastics |> 
  distinct(country) |>
  pull(country)
  
result <- countries |>
  map(.f = Yearly_PET_Prop) |>
  set_names(countries)
```
<details> <summary>Show table output</summary>

```{r}
#| echo: false
result
```

Seems like a significant reduction in Polyester Plastics collected in Argentina. 

## Low & High Density Polyethylene Total Count By Year

*No new columns*

*Requires custom function*

```{r}
sum_pe <- function(input_country){
  plastics |> 
    group_by(year) |>
    filter(country == input_country) |>
    summarize(
      ldpe_total = sum(ldpe, na.rm = TRUE), 
      hdpe_total = sum(hdpe, na.rm = TRUE),                
      .groups = "drop"
      )
}

# Test Case
sum_pe("Brazil")
```

*Iteration*
```{r}
#| echo: true
#| output: false
 
result <- countries |>
  map(.f = sum_pe) |>
  set_names(countries)

```
<details> <summary>Show table output</summary>

```{r}
#| echo: false
result
```


